DATA 624 Project 1 Part A - ATM

### 2020-03-14 DATA624/Project 1-A ATM.Rmd
More work on Project 1

```{r setup, echo=FALSE, cache=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment=NA, message=FALSE, warning=FALSE)
library(tidyverse)
library(lubridate)
library(kableExtra)
library(readxl)
library(fpp2)
library(prophet)
library(forcats)
```

@@

atm %>%
  ggplot(aes(DATE, Cash, color=ATM)) +
  geom_line(lwd=1) +
  ggtitle("Cash Withdrawal by Date and ATM") +
  scale_color_brewer(palette = "Set1") + 
  theme(legend.title = element_blank(),
        legend.position = "bottom",
        axis.title = element_blank())
```

@@

```{r}
atm %>%
  filter(ATM != "ATM4") %>%
  filter(!is.na(ATM)) %>%
  ggplot(aes(DATE, Cash, color=ATM)) +
  geom_line(lwd=1) +
  ggtitle("Cash Withdrawal by Date and ATM") +
  scale_color_brewer(palette = "Set1") + 
  theme(legend.title = element_blank(),
        legend.position = "bottom",
        axis.title = element_blank())
```

@@

df_to_ts <- function(df){
  ### Converts ATM data frames into ts objects
  # Args:
  #     df (data.frame): The ATM data frame 
  # Returns:
  #     ts (Time-Series): An ATM time series object
  ts(df$Cash, start = decimal_date(as.Date(min(df$DATE))), frequency = 365)
}

@@

  # Split the data frame into training, eval and test sets
  test <- tail(df, n_in_test)
  train <- head(df, -n_in_test)
  eval_ts <- tail(train, n_in_test)
  short_train <- head(df, -(2 * n_in_test))

  # Tranform into ts objects
  if(return_ts){
    train <- df_to_ts(train)
    test <- df_to_ts(test)
    eval_ts <- df_to_ts(eval_ts)
    short_train <- df_to_ts(short_train)
  }

  # Return the data to the user
  train_eval_test_ts <- list(train = train, test = test, evaluation = eval_ts, short_train = short_train)

@@

@@

```{r}
atm_1_mean <- mean(atm_1_df$train)
atm_1_last_observation <- tail(atm_1_df$train, 1)$Cash
atm_1_mean <- mean(atm_1_df$train$Cash)
atm_1_results <- 
  atm_1_prophet$forecast %>%
  mutate(DATE = as.Date(ds), prophet_yhat = yhat) %>%
  select(DATE, prophet_yhat) %>%
  mutate(naive_yhat = atm_1_last_observation, 
         mean_yhat = atm_1_mean) %>%
  merge(atm_1_df$evaluation) %>%
  rename(actual = Cash)
atm_1_results$holt <- tail(atm_1_holt$fitted, forecast_window)
```

>>>>

```{r setup, echo=FALSE, cache=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment=NA, message=FALSE, warning=FALSE)
library(tidyverse)
library(lubridate)
library(kableExtra)
library(readxl)
library(fpp2)
library(prophet)
library(forcats)
options(scipen=999)
```

@@

atm %>%
  ggplot(aes(DATE, Cash, color=ATM)) +
  geom_line(lwd=1) +
  ggtitle("Cash Withdrawal by Date and ATM") +
  scale_color_brewer(palette = "Set1") +
  facet_wrap(.~ATM, ncol = 2) +
  theme(legend.title = element_blank(),
        legend.position = "bottom",
        axis.title = element_blank())
```

@@

```{r}
atm %>%
  filter(ATM != "ATM4") %>%
  filter(!is.na(ATM)) %>%
  ggplot(aes(DATE, Cash, color=ATM)) +
  geom_line(lwd=1) +
  facet_wrap(.~ATM, ncol = 2) +
  ggtitle("Cash Withdrawal by Date and ATM") +
  scale_color_brewer(palette = "Set1") + 
  theme(legend.title = element_blank(),
        legend.position = "bottom",
        axis.title = element_blank())
```

@@

df_to_ts <- function(df){
  ### Converts ATM data frames into ts objects
  # Args:
  #     df (data.frame): The ATM data frame 
  # Returns:
  #     ts (Time-Series): An ATM time series object
  ts(df$Cash, start = decimal_date(as.Date(min(df$DATE))), frequency = 7)
  #msts(df$Cash, seasonal.periods=c(7, 365.25))
}

@@

  # Split the data frame into training, eval and test sets
  full <- df
  test <- tail(df, n_in_test)
  train <- head(df, -n_in_test)
  eval_ts <- tail(train, n_in_test)
  short_train <- head(df, -(2 * n_in_test))

  # Tranform into ts objects
  if(return_ts){
    full <- df_to_ts(full)
    train <- df_to_ts(train)
    test <- df_to_ts(test)
    eval_ts <- df_to_ts(eval_ts)
    short_train <- df_to_ts(short_train)
  }

  # Return the data to the user
  train_eval_test_ts <- list(train = train, test = test, evaluation = eval_ts, short_train = short_train, full = full)

@@

```{r}
autoplot(hw(atm_1_ts$train, h=forecast_window))
```

@@

```{r}
atm_1_mean <- mean(atm_1_df$train)
atm_1_last_observation <- tail(atm_1_df$train, 1)$Cash
atm_1_mean <- mean(atm_1_df$train$Cash)
atm_1_results <- 
  atm_1_prophet$forecast %>%
  mutate(DATE = as.Date(ds), prophet_yhat = yhat) %>%
  select(DATE, prophet_yhat) %>%
  mutate(naive_yhat = atm_1_last_observation, 
         mean_yhat = atm_1_mean) %>%
  merge(atm_1_df$evaluation) %>%
  rename(actual = Cash)
atm_1_results$holt <- tail(atm_1_holt$fitted, forecast_window)
```
```{r}
day_of_week_averages <- atm_1_df$train %>%
  mutate(day_of_week = wday(DATE)) %>%
  group_by(day_of_week) %>%
  summarise(yhat = median(Cash))
temp <- atm_1_df$test %>%
  mutate(day_of_week = wday(DATE)) %>%
  merge(day_of_week_averages) %>%
  arrange(DATE) %>%
  select(DATE, yhat)
atm_1_results$day_of_week_model <- temp$yhat
```

### Random Forest

```{r}
library(randomForest)
set.seed(42)
rf_data <- function(df){
  df %>%
    mutate(day_of_week = as.factor(wday(DATE)),
           DATE = as.numeric(DATE))
}

rf_fit <- randomForest(Cash~., rf_data(atm_1_df$train))
predict(rf_fit, rf_data(atm_1_df$test))

atm_1_results$rf <- predict(rf_fit, rf_data(atm_1_df$test))
```

```{r}
lm_fit <- lm(Cash~day_of_week, rf_data(atm_1_df$train))
atm_1_results$lm <- predict(lm_fit, rf_data(atm_1_df$test))
```
